
# Data Preparation I

If you have imported a not previously edited output file from the server you ran the experiment on - you now should be confronted with a mess of columns and rows.


## Select Relevant Variables
A first recommended step would be to subset all relevant columns and get rid of the irrelevant ones. You could either browse through the dataframe or run functions like `str(dat)` or `head(dat)` to get an idea of what variables might be important.

This step could look something like the following code:
```{r, echo = TRUE}

dat <- read.table("myfile.csv", header = TRUE, sep = ";")

# subset relevant variables
dat2 <- dat[, c("Subject", "Block", "Running.LogLevel5.", "Colour", "Flanker",
                "POS", "Target", "ITIDuration", "COMP", "Target.RT",
                "Target.ACC")]

# rename relevant variables
names(dat2) <- c("Subject", "Block", "Task", "Colour", "Flanker", "Position",
                 "Target", "ITIDuration", "Congruency", "RT", "ACC")

```


## Create New Variables

You most likely need to create new variables also. Depending on your research question, this could be for example consecutive trial numbers for each block, 



```{r, echo = TRUE}

# consecutive trial number in each block
dat2$BlockTrialNr <- with(dat2, ave(seq_along(Block), Block + (100 * Subject), FUN = seq_along))

# 



```




# Data Preparation II
\newpage

# References



renaming: if you didnt create the output file by yourself (e.g., assigned variable values and names which should be stored, it can sometimes be a pain to find the correct variables, for example Congruency is often named Cong or Comp and numerically coded (e.g., with values 1 and 2)

# Data preparation

A few examples for new variables with simple ifelse() statements or dyplr functions

a) different task types (e.g., flanker and Simon or different proportions of )
```{r, }
# assign "Flanker" to each value which starts with "Flanker*", else assign value "Simon"
dat2$Task_Type <- ifelse(dat2$Task == "FlankerList20"
                         | dat2$Task == "FlankerList50"
                         | dat2$Task == "FlankerList80", "Flanker", "Simon")
```

b) different proportions of congruent and incongruent trials. You could also use dyplr functions
```{r, }
# assign "Flanker" to each value which starts with "Flanker*", else assign value "Simon"

dat2$Task_Type <- dyplr::startsWith(...)
```


c) task switching
...

d) N-1 variables: if you want to investigate in congruency sequence effects (CSE), or post-error adjustments, you need variables 

```{r, }
# the function lag from the dplyr package gives the value of the previous trial value of a defined variable. You can either call the function by loading the package (library(dplyr)) - or simply call it without loading the package dplyr::lag(...)

# Accuracy N-1
dat2$ACC_N_1 <- dplyr::lag(dat2$ACC)

# Congruency N-1
dat2$Congruency_N_1 <- dplyr::lag(dat2$Congruency)

```


## data types
integer vs numeric?

factorize variables (get correct value for congruent vs incongruent)





## dplyr

stacking commands, few code lines, not always easy to understand or to 

(tidyverse book link)
```{r, }
dat2 <- dat %>%
  mutate(
  Task = Running.LogLevel5.,
  Task_Type = as.factor(case_when(
    startsWith(Task, "FlankerList") ~ "Flanker",
    startsWith(Task, "SimonList")~ "Simon")),
  Position = POS,
  Congruency = as.factor(case_when(
           COMP == 1 ~ "congruent",
           COMP == -1 ~ "incongruent")),
  RT = Target.RT,
  ACC = Target.ACC,
  Congruency_N_1 = lag(Congruency),
  ACC_N_1 = lag(ACC)) %>%
    group_by(Block, Subject) %>%
    mutate(BlockTrialNr = row_number()) %>%
    ungroup %>%
    select(Subject, Block, Task, Task_Type, Colour, Flanker, Position,
  Target, ITIDuration, Congruency, Congruency_N_1, BlockTrialNr, RT, ACC,
  ACC_N_1)
```


